---
title: "Homework 2"
---


```{r}
Traveldata <- read.csv("Sept11Travel.csv")
Traveldata
```

```{r}
library(forecast)
set.seed(2018)
```


1. Plot the pre-event AIR time series. What time series components appear?
#Answer: There is an upward linear trend and seasonality. Solution includes plots for rail and vehicle becauseof question 6.
Create a shorter time series(2001, 8)
```{r}
#pre-event time series for Air 
Air.ts <- ts(Traveldata$Air,  start = c(1990, 1), end = c(2001, 8), frequency = 12)
Air.ts
Air.plot <- plot(Air.ts, ylab = "Air Passenger Miles(Millions)", type="l", bty="l",  main = "Airline Revenue Passenger Miles pre 9/11")

#pre-event time series for Rail
Rail.ts <- ts(Traveldata$Rail,  start = c(1990, 1), end = c(2001, 8), frequency = 12)
Rail.ts
Rail.plot <- plot(Rail.ts, ylab = "Rail Passenger Miles (Millions)", type="l", bty="l", main = "Rail Passenger Miles before 9/11")

# Pre-event time series for Car
vehicle.ts<- ts(Traveldata$VMT, start = c(1990, 1), end = c(2001,8), frequency = 12)
VMTPlot<- plot(vehicle.ts, ylab = "Car Passenger Miles", type="l", bty="l", main = "Vehicle Passenger Miles pre 9/11")  
```

2. The following figure shows a time plot of the seasonally adjusted pre-September-11 AIR series. Which of the following methods would be adequate for forecasting the series shown in the figure?
#Answer: Linear Regression Model with trend and seasonality

        
        

3. Specify a linear regression model for the AIR series that would produce a seasonally adjusted series similar to the one shown in the figure below, with multiplicative seasonality. What is the outcome variable? What are the predictors?
#Answer: Outcome variable: Seasonally adjusted Time 
#        Predictors:  Trend and season

```{r}
# split the data into training and validation
# suppress seasonality using the decompose function

require(xts)
library(scales)
Decompose_AirMultiplicative.ts=decompose(Air.ts,"multiplicative")
Adjusted_Air_multiplicative.ts <- Air.ts-Decompose_AirMultiplicative.ts$seasonal  #With Adjusted seasonality
#PCP <- ts(Adjusted_Air.ts/1000000, start = c(1991,1), end =c(2001,09), frequency = 12)
plot(Adjusted_Air_multiplicative.ts/1000000) 

Decompose_Air_Additive.ts=decompose(Air.ts,"additive")
Adjusted_Air_Additive.ts <- Air.ts-Decompose_Air_Additive.ts$seasonal  #With Adjusted seasonality
#PCP <- ts(Adjusted_Air.ts/1000000, start = c(1991,1), end =c(2001,09), frequency = 12)
plot(Adjusted_Air_Additive.ts/1000000) 

# Using  Decompose function to suppress Seasonality for Rail data
Decompose_Rail.ts=decompose(Rail.ts,"additive")
Adjusted_Rail.ts <- Rail.ts-Decompose_Rail.ts$seasonal  #With Adjusted seasonality
plot(Adjusted_Rail.ts/1000000) 

# Using  Decompose function to suppress Seasonality for Vehicle data
Decompose_vehicle.ts=decompose(vehicle.ts,"additive")
Adjusted_vehicle.ts <- vehicle.ts-Decompose_vehicle.ts$seasonal  #With Adjusted seasonality
plot(Adjusted_vehicle.ts) 

```


4. Run the regression model from (3). Remember to use only pre-event data.
       What can we learn from the statistical insignificance of the coefficients for October and September?
#the statistical insignificance of the coefficients indicate that autocorrelation exists in the dataset and that the data doesnt consider october and sept 

The residual is 1517000 


```{r}
#Adding trend or season or Trend+season
AirMulti.lm <- tslm(Adjusted_Air_multiplicative.ts ~ trend)
summary(AirMulti.lm)

AirMulti.lm.season <- tslm(Adjusted_Air_multiplicative.ts ~ season)
summary(AirMulti.lm.season)

AirMulti.lm.seasonAndTrend <- tslm(Adjusted_Air_multiplicative.ts ~ trend+season)
summary(AirMulti.lm.seasonAndTrend)
plot(AirMulti.lm.seasonAndTrend$residuals)
AirMulti.lm.seasonAndTrend$residuals
```



5.Create an ACF (autocorrelation) plot of the regression residuals.
    What does the ACF plot tell us about the regression model's forecasts?
#Here, we see that there exists trend component as well as a seasonal component. On making first differentiation, it will change the trend effects and seasonal differencing will automatically control the seasonal effects
     How can this information be used to improve the model?
#Using this information, the model can be improved by removing the variables that have high autocorrelation
```{r}
Acf(AirMulti.lm.seasonAndTrend$residuals, lag.max = 12, main="")
#partial autocorrelation
Pacf(AirMulti.lm.seasonAndTrend$residuals, lag.max = 12, main="")
  
```



6.  Fit linear regression models to Air, Rail, and to Auto with additive seasonality and an appropriate trend. For Air and Rail, fit a linear trend. For Rail, use a quadratic trend. Remember to use only pre-event data. Once the models are estimated, use them to forecast each of the three post-event series.
        For each series (Air, Rail, Auto), plot the complete pre-event and post-event actual series overlayed with the predicted series.


```{r}
#adding linear trend for Air,rail and Vehicle
# adding linear trend for Air,rail and Vehicle
Air.lm <- tslm(Adjusted_Air_Additive.ts ~ trend)
summary(Air.lm)

Rail.lm <- tslm(Adjusted_Rail.ts ~ trend)
summary(Rail.lm)

Vehicle.lm <- tslm(Adjusted_vehicle.ts ~ trend)
summary(Rail.lm)


# fitting Quadratic trend for Rail
Rail.lm <- tslm(Adjusted_Rail.ts ~ trend + I(trend^2))
summary(Rail.lm)

```


```{r}
#  generate the naive and seasonal naive forecasts
Airnaive.pred <- forecast::naive(Adjusted_Air_Additive.ts, h = 36)
Airsnaive.pred <- forecast::snaive(Adjusted_Air_Additive.ts, h = 36)
Airnaive.pred
Airsnaive.pred

Railnaive.pred <- forecast::naive(Adjusted_Rail.ts, h = 36)
Railsnaive.pred <- forecast::snaive(Adjusted_Rail.ts, h = 36)
Railnaive.pred
Railsnaive.pred

Vehiclenaive.pred <- forecast::naive(Adjusted_vehicle.ts, h = 36)
Vehiclesnaive.pred <- forecast::snaive(Adjusted_vehicle.ts, h = 36)
Vehiclenaive.pred
Vehiclesnaive.pred
```


```{r}
# plot forecasts and actuals in the training and validation sets
plot(Air.ts,  ylab = "Air Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "") 
axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Airnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Airsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 700999930)) 
lines(c(2004.25, 2004.25), c(0, 700999930))
text(1994.25, 69103630, "Training")
text(2002.75, 69999990, "Validation")
text(2004.90, 69103630, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)


plot(Rail.ts,  ylab = "Rail Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "") 
axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Railnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Railsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 804013874)) 
lines(c(2004.25, 2004.25), c(0, 804013874))
text(1996, 664013874, "Training")
text(2002.75, 664013874, "Validation")
text(2004.90, 664013874, "Future")
#arrows(2001, 630000000, 1990, 630000000, code = 3, length = 0.1, lwd = 1,angle = 30)
#arrows(2004.5 - 3, 2450, 2004, 2450, code = 3, length = 0.1, lwd = 1,angle = 30)
#arrows(2004.5, 2450, 2006, 2450, code = 3, length = 0.1, lwd = 1, angle = 30)
  
  plot(vehicle.ts,  ylab = "Vehicle Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "") 
  axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
  lines(Vehiclenaive.pred$mean, lwd = 2, col = "blue", lty = 1)
  lines(Vehiclesnaive.pred$mean, lwd = 2, col = "red", lty = 1)
  lines(c(2004.70 - 3, 2004.70 - 3), c(0, 350)) 
  lines(c(2004.25, 2004.25), c(0, 270))
  text(1994.25, 240, "Training")
  text(2002.75, 255, "Validation")
  text(2004.90, 240, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)
```


  What can be said about the effect of the September 11 terrorist attack on the three modes of transportation? Discuss the magnitude of the effect, its time span, and any other relevant aspects.

#It is found that the September 2011 terrorist attack has a great impact on the Air and Rail. The magnitude of the effect is very high and it reduces the significance of the model. Vehicle travel did not show a lot of change after 9/11, predicted and actual graph are almosyt similar. 

