---
Title: "Time Series Assignment 1" 
By: "Mercy Nani"
---


1. Is the goal of this study descriptive or predictive?
The purpose of this study was to provide a better understanding of passenger travel behavior patterns that took place after 9/11. Since it is trying to explain the  patterns that took place the goal of this study is descriptive. 

2. Plot each of the three pre-event time series (Air, Rail, Car)
```{r}
library(forecast)
```


```{r}
Traveldata <- read.csv("Sept11Travel.csv")
Traveldata 
```


```{r}
#pre-event time series for Air 

Air_pre911.ts <- ts(Traveldata$Air,  start = c(1990, 1), end = c(2001, 8), frequency = 12)
Air_pre911.plot <- plot(Air_pre911.ts, ylab = "Air Passenger Miles(Millions)", type="l", bty="l",  main = "Airline Revenue Passenger Miles before 9/11")


```
```{r}
#pre-event time series for Rail

Rail.ts <- ts(Traveldata$Rail,  start = c(1990, 1), end = c(2001, 8), frequency = 12)
Rail.plot <- plot(Rail.ts, ylab = "Rail Passenger Miles (Millions)", type="l", bty="l", main = "Rail Passenger Miles before 9/11")


```


```{r}
# Pre-event time series for Car
vehicle.ts<- ts(Traveldata$VMT, start = c(1990, 1), end = c(2001,8), frequency = 12)
VMTPlot<- plot(vehicle.ts, ylab = "Car Passenger Miles", type="l", bty="l", main = "Vehicle Passenger Miles pre 9/11")  
  
```
1) what time series components appear from the plot?

Time series components appear in all three plots. The air and car plots appear to have an upward linear trend, while the rail plots appear to have an exponential trend.
These plots demonstrate all three time series components: Level, Trend, and Seasonality.Seasonality is evident when zooming and in the season plots (in the plots below).



Air Plots: An upward, linear trend and seasonality. 
```{r}
airlinear <- tslm(Air_pre911.ts   ~ trend)
plot(Air_pre911.ts , xlab = "Year", ylab = "Airline Revenue Passenger Miles", bty = "l", main = "Trend: Air Revenue Passenger Miles")
lines(airlinear$fitted, lwd=2)
```


```{r}
yearlyAir <- aggregate(Air_pre911.ts, nfrequency=1, FUN=sum)
plot(yearlyAir, xlab = "Year", ylab = "Air Passenger Miles",bty="l", main = "Yearly Air Miles with Seasonality Suppressed")
```

```{r}
#zooming in 
air.ts.zoom <- window(Air_pre911.ts, start = c(1999, 1), end = c(2000, 12))
plot(air.ts.zoom, xlab = "Year", ylab = "Airline Revenue Passenger Miles", bty = "l", main = "Zoom: Air Revenue Passenger Miles")
```

```{r}
ggseasonplot(Air_pre911.ts, col =rainbow(12), year.labels=TRUE, ylab = "Air Passenger Revenue Miles", main = "Seasonal Plot for Air Passenger Revenue")
```


```{r}
# Visualizing Air RPM Time Series
air_seasonality <- stl(Air_pre911.ts, s.window = "period")
plot(air_seasonality)
```


Rail Plots 
#shows a downward trend and seasonality
```{r}
#show downward linear trend
Rail.ts <- ts(Traveldata$Rail, start = c(1990, 1), end = c(2001, 8), frequency = 12)
trainlinear <- tslm(Rail.ts ~ trend)
plot(Rail.ts, xlab = "Year", ylab = "Rail Passenger Miles", bty = "l", main = "Trend: Rail Passenger Miles")
lines(trainlinear$fitted, lwd = 2)

```

```{r}
quarterlyRail <- aggregate(Rail.ts, nfrequency=4, FUN=sum)
plot(quarterlyRail, xlab = "Year", ylab = "Rail Passenger Miles", bty="l", main = "Rail Miles by Quarter with Seasonality Suppressed")
```

```{r}
ggseasonplot(Rail.ts, col =rainbow(12), year.labels=TRUE, ylab = "Rail Passenger Revenue Miles", main = "Season Plot for Rail Miles Travelled")
```

```{r}
Rail.ts.zoom <- window(Rail.ts, start = c(1998, 1), end = c(2000, 12))
plot(Rail.ts.zoom, xlab = "Year", ylab = "Rail Passenger Miles", bty = "l", main = "Zoom: Rail Passenger Miles")
```



Car Plots: show an upward trend and seasonality
```{r}
carlinear <- tslm(vehicle.ts ~ trend)
plot(vehicle.ts, xlab = "Year", ylab = "Car Passenger Miles", bty = "l", main = "Trend: Vehicle Passenger Miles")
lines(carlinear$fitted, lwd = 2)
```

```{r}
quarterlyAuto <- aggregate(vehicle.ts, nfrequency=4, FUN=sum)
plot(quarterlyAuto, xlab = "Year", ylab = "Car Passenger Miles",main = "Quarterly Vehicle with Seasonality Suppressed", bty="l")
```

```{r}
yearlyAuto <- aggregate(vehicle.ts, nfrequency=1, FUN=sum)
plot(yearlyAuto, xlab = "Year", ylab = "Car Passenger Miles", main = "Yearly Car with Seasonality Suppressed",  bty="l")
```

```{r}
ggseasonplot(vehicle.ts, col =rainbow(12), year.labels=TRUE, ylab = "Car Passenger Miles", main = "Season Plot for Vehicle Miles Travelled", bty="l")
```

```{r}
car.ts.zoom <- window(vehicle.ts, start = c(1998, 1), end = c(2000, 12))
plot(car.ts.zoom, xlab = "Year", ylab = "Car Passenger Miles", bty = "l", main = "Zoom: Car Passenger Miles")
```

