---
title: "Time Series Assignment 1" 
By: "Mercy Nani"
---


1. Is the goal of this study descriptive or predictive?
The purpose of this study was to provide a better understanding of the patterns that took place after 9/11 so the goal of this study is descriptive. 

2. Plot each of the three pre-event time series (Air, Rail, Car).
What time series components appear from the plot?
What type of trend appears? Change the scale of the series, add trendlines and suppress seasonality to better visualize the trend pattern.
```{r}
library(forecast)
```



```{r}
Traveldata <- read.csv("Sept11Travel.csv")
Traveldata 
```


```{r}
#pre-event time series for Air 

air_pre911.ts <- ts(Traveldata$Air, start = c(1990,1), end = c(2001,8), freq = 12)
plot(air_pre911.ts, xlab = "Time", ylab = "Air.RPM..000s.", ylim = c(29000000, 70000000), bty = "l")


```
```{r}
#pre-even time series for Rail
pre_9_11_rail <- ts(sept_11$Rail, start = c(1990, 1), end = c(2001, 8), freq = 12)
plot(pre_9_11_rail, xlab = "Time", ylab = "Rail PM", ylim = c(300000000, 700000000), bty = "l")
```


```{r}
# Pre-event car time series plot
pre_9_11_car <- ts(sept_11$VMT, start = c(1990, 1), end = c(2001, 8), freq = 12)
plot(pre_9_11_car, xlab = "Time", ylab = "Car VMT", ylim = c(150, 262), bty = "l")
```

Time series components appear in all three plots. The air and car plots appear to have an upward linear trend, while the rail plots appear to have an exponential trend.
These plots demonstrate all three time series components: Level, Trend, and Seasonality.
We see fairly pronounced seasonality with all three time series, but different trend shapes at play:

```{r}
# Visualizing Air RPM Time Series
air_seasonality <- stl(pre_9_11_air, s.window = "period")
plot(air_seasonality)
```


```{r}

```

