---
title: "Segmenting Consumers of Bath Soap"
output:
  html_document: default
  html_notebook: default
---
Load the folowing libraries
```{r}
library(cluster)
library(NbClust)
library(flexclust)
library(fMultivar)
#library(ggplot2)
library(lattice)
library(gridBase)
library(gridExtra)
#install.packages("plyr")
library(plyr)
```


Read CSV file into R
```{r}
BathSoap <- read.csv("C:/Users/Mercy/Downloads/BathSoap.csv")
View(BathSoap)

#Data contains the % symbol, so I removed it in excel to make R read it as a numeric variable
BathSoap1 <- read.csv("C:/Users/Mercy/Downloads/BathSoap1.csv")
View(BathSoap1)
```


```{r}
#Graphical representation of missing values
library(VIM)
aggr(BathSoap1)
```

Data preprocessing
```{r}
#Data has many missing values so imputation of values was done
BathSoap1$SEX[BathSoap1$SEX=="0"]<-2 #converted 0 to 2 to represent "female"
BathSoap1$FEH[BathSoap1$FEH=="0"]<-1 #assuming major population is vegetarian
BathSoap1$MT[BathSoap1$MT=="0"]<-10 #most common language spoken
BathSoap1$HS[BathSoap1$HS=="0"]<-5 #average number of people in household
BathSoap1$EDU[BathSoap1$EDU=="0"]<-5 #majority of student have 12th standard education
BathSoap1$CS[BathSoap1$CS=="0"]<-1 #majority of people have TV at home

#Nullifyin member ID
row.names(BathSoap1)<-BathSoap1$Member.id
BathSoap1$Member.id<-NULL
BathSoap2<-BathSoap1

#scaling data
BathSoap2<-lapply(BathSoap2, scale)
BathSoap2<-as.data.frame(BathSoap2)

#To calculate Brand loyalty Index
BathSoap2$MaxVolScore<-  apply(BathSoap1[,23:30],1,max)
BathSoap2$BrandLoyaltyIndex<-0.4 *BathSoap2$No..of.Brands + 0.4*BathSoap2$Brand.Runs + 0.2*BathSoap2$MaxVolScore
View(BathSoap2)

```



```{r}
#Ward-linking cluster on normalized data

d <- dist(BathSoap2) #Euclidean
summary(d)
fit.w <- hclust(d,method="ward.D")
fit.w$height
fit.w$order
plot(fit.w,hang=-1,cex=0.8,main="Ward linking clusters")

```


```{r}
#Dividing data into variables that describe puchase behaviour and variables that describe the basis for purchase
Purchase_behavior <- BathSoap2[,c(1:10,13:21,46:47)]
Purchase_basis <- BathSoap2[,c(31:45)]
```


K-Means clustering on purchase behavior dataset
```{r}
#install.packages("cluster")
library(NbClust)
set.seed(1234)
nc <- NbClust(data = Purchase_behavior,distance = "euclidean", min.nc=2,max.nc=5,method="kmeans",index = "all", alphaBeale = 0.1)

#Bestnumber of clusters are 5
table(nc$Best.nc[1,])
barplot(table(nc$Best.n[1,]),
        xlab="Number of Clusters", ylab="Number of criteria",
        main = "Number of clusters chosen by Criteria ")


km = kmeans(Purchase_behavior, 5, nstart=25)
km
km$size
dist(km$centers)

#install.packages("factoextra")
library(factoextra)
library(cluster)
#library(ggpubr)
fviz_nbclust(Purchase_behavior, kmeans, method = "wss") + geom_vline(xintercept = 5, linetype = 2)+labs(subtitle = "Elbow method")
wss <- numeric(5)
for (k in 1:5)
  wss[k] <- sum(kmeans(Purchase_behavior, centers=k, nstart=25)$withinss)
plot(1:5,wss,type="b",xlab="Number of Clusters", ylab="Within Sums of squares")
wss[]

```


K-Means for purchase basis dataset
```{r}

nc1 <- NbClust(data = Purchase_basis,distance = "euclidean", min.nc=2,max.nc=5,method="kmeans",index = "all", alphaBeale = 0.1) 

#Bestnumber of clusters are 2
table(nc1$Best.nc[1,])
barplot(table(nc1$Best.n[1,]),
        xlab="Number of Clusters", ylab="Number of criteria",
        main = "Number of clusters chosen by Criteria ")
km1 = kmeans(Purchase_basis, 2, nstart=25)
km1
km1$size
dist(km1$centers)
fviz_nbclust(Purchase_basis, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Elbow method")

```


K-Means for purchase bahavior and purchase basis datasets
```{r}
nc2 <- NbClust(data = BathSoap2[,c(1:10,13:21,31:45,47)],distance = "euclidean", min.nc=2,max.nc=5,method="kmeans",index = "all", alphaBeale = 0.1) 

#Bestnumber of clusters are 2
table(nc2$Best.nc[1,])
barplot(table(nc2$Best.n[1,]),
        xlab="Number of Clusters", ylab="Number of criteria",
        main = "Number of clusters chosen by Criteria ")
km2 = kmeans(BathSoap2[,c(1:10,13:21,31:45,47)], 2, nstart=25)
km2
km2$size
dist(km2$centers)

fviz_nbclust(BathSoap2[,c(1:10,13:21,31:45,47)], kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Elbow method")


```


Question 2
To select the best segmentation we have to see the distance between clusters
looking at all three data sets 
```{r}

p1 <- fviz_cluster(km, geom = "point",   data=Purchase_behavior) + ggtitle("k = 5")
p2 <- fviz_cluster(km1, geom = "point",  data = Purchase_basis) + ggtitle("k = 2")
p3 <- fviz_cluster(km2, geom = "point",  data = BathSoap2[,c(1:10,13:21,31:45,47)]) + ggtitle("k=2 combined data")

library(gridExtra)
grid.arrange(p1, p2, p3, nrow = 2)
BathSoap3<-BathSoap2[,c(1:10,13:21,31:45,47)]



```

Based on this we see that the cluster with the combined data gives the higest degree of separation between clusters making it the best segmentation criteria.
characteristics of the thrid cluster

The combined cluster dataset is best to create clusters because there is a clearly clearly defined separation between clusters.It also considers most of the variables.


```{r}
#Model Prediction using the best cluster on a sample test dataset (first 50 rows).
#install.packages("FNN")
#install.packages("flexclust")
library(FNN)
library(flexclust)
testdata<-BathSoap3[1:50,]
pred.knn <- get.knnx(km2$centers, testdata, 1)$nn.index[,1]
pred.knn

```
